Timbral similarity between extended playing techniques: the next frontier in querying instrumental databases

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on Hornbostel-Sachs
trivial commits

\newcommand{\ml}[1]{\textcolor{blue}{ML : #1}}
\newcommand{\vl}[1]{\textcolor{purple}{VL : #1}}

@@

\vl{On pairwise similarity judgments of timbre, cite:
McAdams, S et al. (1995). Perceptual scaling of synthesized musical
timbres: common dimensions, specificities, and latent subject
classes. In: Psychological research 58.3, pp. 177?192}

\vl{On the musicology of extended playing techniques, cite:
Kostka, S (2016). Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11. #biblio}

\vl{On the two definitions of timbre, cite:
 Castellengo, M and D Dubois (2007). Timbre ou timbres? Propriété
du signal, de l?instrument, ou construction cognitive? In: Cahiers
de la Société québécoise de recherche en musique 9, pp. 25?38
#biblio}

@@

As explained in the first section, gathering perceptual similarity data is plagued is plagued with a size problem. Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte. In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound examplar we consider the whole range of nuance and pitch that is available. This leads to a dataset of 9346 sound recordings. It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken in this paper as a reasonable assumption.
\vl{It's too bad that this "reasonable" assumption is not challenged against the data. In music psychology, the "bandwidth for timbre invariance" is of the order of one octave (Handel and Erickson, Music Perception, 2001) or less (Steele and Williams, Music Perception, 2006).
One way to do it without collecting new similarity judgments would be to break down the precision by pitch of query.
Presumably we have an inverted U shape with a peak near C4.
#research #figure #biblio}

>>>>

\newcommand{\ml}[1]{\textcolor{blue}{ML : #1}}
\newcommand{\vl}[1]{\textcolor{red}{VL : #1}}

@@

\vl{On pairwise similarity judgments of timbre, cite:
McAdams, S et al. (1995). Perceptual scaling of synthesized musical
timbres: common dimensions, specificities, and latent subject
classes. In: Psychological research 58.3, pp. 177-192}

\vl{On the musicology of extended playing techniques, cite:
Kostka, S (2016). Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11. @biblio}

\vl{On the two definitions of timbre, cite:
 Castellengo, M and D Dubois (2007). Timbre ou timbres ? Propri\'{e}t\'{e}
du signal, de l'instrument, ou construction cognitive ? In: Cahiers
de la Soci\'{e}t\'{e} qu\'{e}b\'{e}coise de recherche en musique 9, pp. 25-38
 @biblio}

\vl{We should probably say a word on the Hornbostel-Sachs taxonomy,
which is based on player-instrument interaction, but is not outdated
given the variety of contemporary music techniques. @biblio}

@@

As explained in the first section, gathering perceptual similarity data is plagued is plagued with a size problem. Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte. In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound examplar we consider the whole range of nuance and pitch that is available. This leads to a dataset of 9346 sound recordings. It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken in this paper as a reasonable assumption.
\vl{It's too bad that this "reasonable" assumption is not challenged against the data. In music psychology, the "bandwidth for timbre invariance" is of the order of one octave (Handel and Erickson, Music Perception, 2001) or less (Steele and Williams, Music Perception, 2006).
One way to do it without collecting new similarity judgments would be to break down the precision by pitch of query.
Presumably we have an inverted U shape with a peak near C4.
@research @figure @biblio}
